---
title: "Bonanza Creek Snowshoe Hare Report"
author: "Conner Smith"
date: "11/3/2021"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages. Hiding the code chunk from the knitted document. 

library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(kableExtra)
library(broom)
library(effsize)
library(equatiomatic)
```

# **1. Introduction**

This study explores population dynamics and characteristics of Snowshoe Hares measured at the Bonanza Creek Experimental Forest Long Term Ecological Research site. The analysis looks at the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012 (Keilland et al. 2017). Hares were trapped across three different experimental sites featuring a range of different ecological characteristics. This report looks specifically at the number of juvenile hares trapped over the study period and compares the relationship between study variable including weight, sex, and hind foot length. The analysis includes several tests to classify differences between subsets in this sample. 

```{r, include = FALSE}
# Read in the snowshoe hare data, hiding the code chink from the knitted document. 

hares <- read_csv(here("data", "bonanza_hares.csv"))
  
```

# **2. Data and Methods**

This data was made available by researchers working at the Bonanza Creek LTER, a partnership between the University of Alaska Fairbanks, and the U.S. Forest Service (Keilland et al. 2017). The data contain 3,197 observations recorded between 1998 and 2012. These data are visualized throughout the report and weights for juvenile hares are compared for both sexes using a two-sample t-tests using a significance level ($\alpha$) of 0.05 throughout. Differences between groups are described by Cohen's *d* effect size. The relationship between hare weight and hind foot length is explored using a simple linear regression. All analyses are in R version 4.1.1 using RStudio version 2021.09.1+372.  

# **3. Juvenile Hare Trapping Summary**

This section includes a summary of the juvenile hare trappings by year. Trappings varied significantly over the study period and are visualized in **Figure 1** below. 

#### **Figure 1: Count of Bonanza LTER Juvenile Snowshoe Hare Trappings by Year (1998-2012)**

```{r}
# Filter the data to account for only juvenile hares and use `lubridate:: mdy() to clean the dates to allow year to be separated. Then Count the total number of juvenile hare trappings `count()` during each year of the study with `group_by()`.

hares_juvenile <- hares %>% 
  filter(age == "j") %>% 
  mutate(date_new = mdy(date)) %>%
  mutate(year = year(date_new)) %>% 
  group_by(year)

juvenile_summary <- hares_juvenile %>% 
  count(year) %>% 
  rename(trappings = "n")

# Create a finalized data visualization of the counts by year with `geom_histogram()`.
  
ggplot(data = hares_juvenile, aes(x = year)) +
  geom_histogram(fill = "slategray2",
                 show.legend = FALSE, binwidth = 0.5) +
  theme_bw() +
  scale_x_continuous(breaks = 1998:2012)+
  labs(x = "Year", y = "Count of Trappings")
```


```{r, include=FALSE}
# Including calculation of mean/median values for the description without generating a summary table. Hide from knitted document. 

juvenile_mean <- mean(juvenile_summary$trappings)
juvenile_median <- median(juvenile_summary$trappings)
juvenile_max <- max(juvenile_summary$trappings)
juvenile_min <- min(juvenile_summary$trappings)

```


##### **Figure 1:** *This graph shows the spread of juvenile Snowshoe Hare trappings at the Bonanza LTER between 1998 and 2012. There were no trappings recorded in 1998, 2002, and 2009. Data: Keilland et al. (2017).*

The fewest number of juvenile hares (`r juvenile_min`) were trapped in 2010 and the highest number (`r juvenile_max`) were trapped in 1999. The average annual hare trappings is around `r round(juvenile_mean)` across the 12 years where juvenile hares were trapped. The median value is `r juvenile_median`, which falls across two years (2003 and 2011). The figure shows an overall decline in the number of juvenile hares trapped at the research site. In order to more confidently assert that this represents an actual decline in the juvenile hare population, the number of traps and time spent trapping would have to be standardized across all study years and sites. If this was not possible for some reason, researchers could use the existing data and create weightings based on number of traps and time spent so years where trapping activity was less are not leading to skewed results. 

# **4. Juvenile Hare Weight Analysis by Sex and Site**

This section looks for differences in juvenile hare wights across the three study sites and compares the differences by sex. It includes a summary figure, table, and analysis using an unpaired, two-sample t-test. 

#### **Figure 2: Weight Distribution of Bonanza LTER Juvenile Snowshoe Hares by Site and Sex**

```{r}
# Compare juvenile hare weights by sex and site (grid), include entries where sex was not recorded (`NA`). Using a jitter plot to show these comparisons. 

juvenile_weight <- hares_juvenile %>% 
  select(grid, sex, weight) %>% 
  mutate(grid = case_when(grid == "bonbs" ~ "Lowland Black Spruce",
                               grid == "bonmat" ~ "Bonanza Mature",
                               grid == "bonrip" ~ "Bonanza Riparian")) %>%
  mutate(sex = case_when(sex == "f" ~ "Female",
                               sex == "m" ~ "Male"))

ggplot(data = juvenile_weight, aes(x = sex, y = weight)) +
  geom_jitter(aes(color = sex), width = 0.2, show.legend = FALSE) +
  facet_wrap(~grid) +
  scale_color_manual(values = c("skyblue3", 
                               "slateblue3", 
                               "slategray3")) +
  theme_bw() +
  labs(x = "Sex", y = "Weight (grams)")
```

##### **Figure 2:** *This graph shows the weight distribution of Snowshoe Hare trappings at different site types broken out by hare sex. Entries where sex was not recorded are included as "NA". Data: Keilland et al. (2017).*

**Figure 2** shows that more hares were trapped at the Bonanza Riparian site compared to the Bonanza Black Spruce and Mature sites. It also shows that  the weight differences between juvenile males and female appears to favor males, but the overall difference is minimal. **Table 1** below provides summary statistics of this data . 

**Table 1: Summary Statistics of Bonanza Hare Weight by Sex.** *This table shows summary statistics for hare weights based on trap site and sex. It includes the mean and median values for "NA" entries. Data: Keilland et al. (2017).*

```{r}
# Creating a summary table comparing juvenile weights by sex using `kable()`. 

juvenile_stats <- juvenile_weight %>% 
  group_by(sex) %>% 
  summarize(trappings = n(),
            mean_weight = mean(weight, na.rm = TRUE),
            median_weight = median(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm = TRUE)) 

juvenile_stats %>% 
kable(col.names = c("Sex",
                    "Trappings",
                    "Mean Weights (g)",
                    "Median Size (g)",
                    "Standard Deviation (g)"), digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

```{r, include = FALSE}
# Including some exploration of the differences in data including effect size (Cohen's D) and an unpaired, two-sample t-test. Not including this chunk in the knitted document. 

mean_diff <- juvenile_stats[2, "mean_weight"] - juvenile_stats[1, "mean_weight"] 

mean_percent <- mean_diff / juvenile_stats[2, "mean_weight"]

combined_sample <- juvenile_stats$trappings[1] + juvenile_stats$trappings[2]

# Include exploratory QQ plots to determine if t-tests are appropriate for this data.

ggplot(data = juvenile_weight, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~ sex)

ggplot(data = juvenile_weight, aes(x = weight)) +
  geom_histogram() +
  facet_wrap(~ sex)

# There is no substantial indication in these graphs that leads me to think that a t-test is not appropriate. The central limit theorem can help explain this. 

```


```{r}
# Run the t test and Cohen's d, including this as a code fold in the knitted document. 

juvenile_male <- juvenile_weight %>% 
  filter(sex == "Male") %>% 
  pull(weight)

juvenile_female <- juvenile_weight %>% 
  filter(sex == "Female") %>% 
  pull(weight)

hare_t <- t.test(juvenile_male, juvenile_female, na.rm = TRUE)

hare_cohen <- cohen.d(juvenile_male, juvenile_female, na.rm = TRUE)
```

A closer look at the hare weight distribution by sex does not provide sufficient indication to conclude that the data are not normally distributed. This is in part due to the relatively large sample size (`r combined_sample`). Weights for juvenile male hares do not appear to follow a normal distribution closely, but this could be explained by a limited sample resulting from inconsistencies in trapping and study design.

The mean weight for juvenile male hares (`r round(juvenile_stats$mean_weight[2], 2)` $\pm$ `r round(juvenile_stats$sd_weight[2], 2)` grams, *n* = `r juvenile_stats$trappings[2]`; mean $\pm$ 1 standard deviation) is slightly larger than that for females (`r round(juvenile_stats$mean_weight[1], 2)` $\pm$ `r round(juvenile_stats$sd_weight[1], 2)` grams, *n* = `r juvenile_stats$trappings[1]`; mean $\pm$ 1 standard deviation). 

This gives an absolute difference in means between the juvenile male and female rabbits of `r round(mean_diff, 2)` grams (or a `r round(mean_percent,3)*100` percent difference). The difference is mean weight between the male and female hares included in this sample is significant (two-sample t-test: t(`r round(hare_t$parameter,2)`= `r round(hare_t$statistic,2)`, p = `r round(hare_t$p.value,3)`). However, the effect size is relatively small (Cohen's *d* = `r round(hare_cohen$estimate,2)`). 

# **5. Juvenile Hare Weight and Foot Comparisons** 

This section explores the relationship between juvenile hare wight and foot lengths. It includes a simple linear regression and analysis of whether this is an appropriate method to use for these data. 

#### **Figure 3: Weight and Hind Foot Length of Bonanza LTER Juvenile Snowshoe Hares**

```{r}
# Explore the relationship between juvenile snowshoe hare hind foot length (variable `hindft`, recorded in millimeters), and weight.

hare_foot <- hares_juvenile %>% 
  select(hindft, weight)

ggplot(data = hare_foot, aes(x = weight, y = hindft)) +
  geom_point(color = "slategray3") +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Weight (g)", y = "Hind Foot Length (mm)")
  
```

```{r, include=FALSE}
# Generating clean data frames showing the coefficients of this model, as well as the r squared and Pearson's r value. Not including the code chunk in the knitted document. 

hares_lm <- lm(weight ~ hindft, data = hare_foot)

hares_lm_tidy <- tidy(hares_lm)

hares_lm_glance <- glance(hares_lm)

hares_pearson <- cor.test(hare_foot$weight, hare_foot$hindft, method = "pearson", use = "complete.obs")
```

##### **Figure 3:** *This graph shows the relationship between weight (g) and hind foot length (mm) of snowshoe hares. The graph includes a simple line of best fit assuming a linear relationship between weight and foot length* ($\beta$~1~ = `r round(hares_lm_tidy$estimate[2], 2)` g mm^-1^, p <0.001, $R^2$ = `r round(hares_lm_glance$r.squared[1], 2)`, Pearson's *r* = `r round(hares_pearson$estimate, 2)`). *Data: Keilland et al. (2017).*

The simple linear regression included in **Figure 3** shows that these data may not be best described by a linear relationship. Other variable (such as sex) should also be included in the development of predictive models. The following equation shows the equation associated with this simple linear model:
```{r}
# Use `equatiomatic:: extract()` to generate an equation. 

extract_eq(model = hares_lm, use_coefs = TRUE)
```

Simple linear regression shows that the relationship between hare weight and hind foot length is not completely linear. The model predicts an increase of `r round(hares_lm$coefficients[2], 2)` millimeter increase in hare hind foot length for every gram increase in weight. This model does not significantly predict hare hind foot length from weight (p <0.001, $R^2$ = `r round(hares_lm_glance$r.squared[1], 2)`). This means that simple linear regression only predicts `r round(hares_lm_glance$r.squared[1], 3)*100` percent of the variance in the dependent variable (hind foot length) from the independent variable (weight). Hare weight and hind foot length are significantly, but not strongly correlated (Pearson's *r* = `r round(hares_pearson$estimate, 2)`, p < 0.001). 

This provides further indication of earlier concerns that a simple linear regression does not provide a complete explain the relationship between hare weight and hind foot length. Looking at the shape of the data in **Figure 3** above, it is reasonable to think that a different model (i.e. one with a slightly logarithmic curve) or a model that incorporates more variables could more accurately capture the relationship between Snowshoe Hare weight and foot length. 

# **6. Summary**

The key takeaways from this analysis can be summarized as follows: 

- There is some inconsistency in the number of juvenile hares trapped over the duration of the study with some years (1998, 2002, and 2009) having no trappings. 

- The most juvenile hares were trapped at the Bonanza Riparian site. 

- The actual difference in mean weights between male and female juvenile hares is `r round(mean_diff, 2)` grams. This difference is significant, although the effect size is not large (Cohen's *d* = `r round(hare_cohen$estimate,2)`). 

- The relationship between hare weight and hindfoot length is not best described by a simple linear regression. Under a linear regression, only `r round(hares_lm_glance[1, "r.squared"], 3)*100` percent of the variance in hind foot length is predictable from hare weight. Hare weight and hind foot length are significantly, but not strongly correlated (Pearson's *r* = `r round(hares_pearson$estimate, 2)`. This indicates that a different type of model is likely more appropriate to understand this relationship. 

Possible next steps in this analysis could be:

- Standardizing the inconsistencies in the data across study years. 

- Generating a new model that explains a higher percentage of the variance in hind foot length as a function of weight. 

# **7. Citations**

- Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017.        Snowshoe hare physical data in Bonanza Creek Experimental Forest:           1999-Present ver 22. Environmental Data Initiative.                         https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14*


```{r, include = FALSE}
citation(package = "tidyverse")
citation(package = "here")
citation(package = "lubridate")
citation(package = "janitor")
citation(package = "kableExtra")
citation(package = "broom")
citation(package = "effsize")
citation(package = "equatiomatic")

```


- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source    Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

- Kirill Müller (2020). here: A Simpler Way to Find Your
  Files. R package version 1.0.1.
  https://CRAN.R-project.org/package=here
  
- Garrett Grolemund, Hadley Wickham (2011). Dates and Times
  Made Easy with lubridate. Journal of Statistical Software,
  40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.
  
- Sam Firke (2021). janitor: Simple Tools for Examining and
  Cleaning Dirty Data. R package version 2.1.0.
  https://CRAN.R-project.org/package=janitor
  
- Hao Zhu (2021). kableExtra: Construct Complex Table with
  'kable' and Pipe Syntax. R package version 1.3.4.
  https://CRAN.R-project.org/package=kableExtra
  
- David Robinson, Alex Hayes and Simon Couch (2021). broom:
  Convert Statistical Objects into Tidy Tibbles. R package
  version 0.7.9. https://CRAN.R-project.org/package=broom
  
- Torchiano M (2020). _effsize: Efficient Effect Size
  Computation_. doi: 10.5281/zenodo.1480624 (URL:
  https://doi.org/10.5281/zenodo.1480624), R package version
  0.8.1, <URL: https://CRAN.R-project.org/package=effsize>.
  
- Daniel Anderson, Andrew Heiss and Jay Sumners (2021).
  equatiomatic: Transform Models into 'LaTeX' Equations. R
  package version 0.3.0.
  https://CRAN.R-project.org/package=equatiomatic

