---
title: "Bonanza Creek Snowshoe Hare Report"
author: "Conner Smith"
date: "11/3/2021"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attach packages.
library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(kableExtra)
library(equatiomatic)
```

# **Introduction**

This study explores population dynamics and characteristics of Snowshoe Hares measured at the Bonanza Creek Experimental Forest (Long Term Ecological Research site). The analysis looks at the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012 ... 

*Data citation: Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14*

```{r, include = FALSE}
# Read in the snowshoe hare data 

hares <- read_csv(here("data", "bonanza_hares.csv"))
  
```

# **Data and Methods**

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

## **Initial Investigation**

This section includes a summary of the hare count by year.

#### **Figure 1: Count of Bonanza LTER Snowshoe Hare Trappings by Year**

```{r}
# Count the total number of juvenile hare trappings (`summarize()` during each year of the study (`group_by()`), and create a finalized data visualization of the counts by year.

hares_clean <- hares %>% 
  mutate(date_new = mdy(date)) %>%
  mutate(year = year(date_new)) %>% 
  group_by(year)

hares_summary <- hares_clean %>% 
  count(year) %>% 
  rename(trappings = "n")
  
ggplot(data = hares_clean, aes(x = year)) +
  geom_histogram(fill = "slategray2",
                 show.legend = FALSE) +
  theme_bw() +
  labs(x = "Year", y = "Count of Hare Trappings", 
       title = "Count of Bonanza LTER Snowshoe Hare Trappings (1998-2012)")

# Including calculation of mean/median values for the description without generating a summary table.

mean(hares_summary$trappings)
median(hares_summary$trappings)

```

**Figure 1: Count of Bonanza LTER Snowshoe Hare Trappings by Year.** *This graph shows the spread of Snowshoe Hare trappings at the Bonanza site between 1998 and 2012.*

The fewest number of hares (38) were trapped in 2002 and the highest number (410) were trapped in 2009. The average annual hare trappings is around 213 across the 14 years of data. The median value is 179, which was recorded in 2006. Overall, trappings were highest around 2000 and between 2008 and 2011. 


These are absolute counts, but this count is also impacted by effort (e.g. the number of days and number of traps in the study each year). What would you suggest be done moving forward (you are not expected to actually do this) to standardize the juvenile hare population in future analyses?
